---
title: "ESM 204 Homework 3"
author: "Justine Lang"
date: "2023-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(here)
library(broom)
library(car)
library(pracma)
```

```{r}
### Read in data


```


## Question 1

The data set HW3 data.csv provides price (in $) and quantity (in number of EVs) estimates
of demand per year for both DAC and non-DAC groups. Run linear regressions (with an
intercept) to estimate the demand curves for DAC and non-DAC consumers. What are 2-3
reasons you can think that may explain the differences in demand you estimate?










### Notes

Slope = 50,000/Q_agg(50,000)

P = 0 + 50,000/Q_agg(50,000)

Don't worry about the kink in the agg curve. Just use triangle formula. Write function. Use demand function to calculate base of the triangle. 

Producer surplus simpler. 

Total environmental damage = externality per car x total cars. 

EV tax: the MEC at Qso; in this case, tax just equals MEC

Equate new supply curve (MSC) to aggregate demand curve. P = mQ + t; Q = P-t/m -> supply(p,t); Q should = Qagg. Use uniroot to find Pt*. Plug Pstar back into equation to find Qstar. 

Would work to find Pstar or Qstar first, but might be best to find Pstar first. 

Assume the mining externality is borne entirely by the DAC group. 

Health cost = Qagg(Pt) x MEC

The results should indicate that the tax doesn't do much for the disadvantaged group. 

See tax on sellers graph. 

stargazer

Tax revenue = Qstar x t(MEC); calculate the redistribution four different times.

Just answer question 7 qualitatively. 































